Topology and physical distinguishability

### 2017-07-12 Articles/TopologyAndExperimentalDistinguishability/article.tex
Minor edits of Introduction

We believe we have identified the answer: topologies capture the way objects are experimentally distinguishable. It is open sets that are associated with experimental outcomes, not the values themselves. For example, as real values are only measurable with finite precisions, the standard topology on euclidean spaces is the one that is physically meaningful. Along the same line, experimental relationships will need to preserve what can be experimentally distinguished and they will therefore be continuous functions.

In this sense, experimentally distinguishability is \emph{the} application of topology in experimental science. It is fitting, then, that topology is so wide-spread as defining what can be experimentally identified is truly a fundamental aspect in science. The aim of this paper is to lay down a formal framework that formalizes this insight.

The work is organized as follows. We will first define experimental observations as statements paired with an experimental test that is able to verify them. We will study their properties under logical operations and show they are only closed under finite conjunction and countable disjunction.

We will then show that any set of objects that are experimentally distinguishable is a Hausdorff second-countable topological space. Both the points and the topology of the space will be formally defined in terms of the observations themselves.

Finally, we will study relationships between experimentally distinguishable objects. This will be represented by continuous functions and can be shown to be themselves physically distinguishable

We believe this work not only provides interest insights, but a firm justification of the use of topological ideas in the physical science.

>>>>

We believe we have identified the answer: topologies capture the way objects are experimentally distinguishable. It is open sets that are associated with experimental outcomes, not the values themselves. For example, as real values are only measurable with finite precision, the standard topology on Euclidean spaces is the one that is physically meaningful. Along the same line, experimental relationships need to preserve what can be experimentally distinguished, and they are therefore represented as continuous functions.

In this sense, experimental distinguishability is \emph{the} application of topology in experimental science. It is fitting, then, that topology is so widespread given that defining what can be experimentally identified is truly a fundamental aspect of science. The aim of this paper is to lay down a framework that formalizes this insight.

The work is organized as follows. We first define experimental observations as statements paired with an experimental test that is able to verify them. We study their properties under logical operations and show that they are only closed under finite conjunction and countable disjunction.  We then show that any set of objects that are experimentally distinguishable is a Hausdorff second-countable topological space. Both the points and the topology of the space are formally defined in terms of the observations themselves.  Finally, we study relationships between experimentally distinguishable objects. These are represented by continuous functions, and they can be shown to be themselves physically distinguishable.

We believe this work not only provides interesting insights, but a firm justification of the use of topological ideas in the physical sciences.
